smdi: An R package to perform structural missing data investigations on partially observed confounders in real-world evidence studies

### 2024-01-02 manuscript/manuscript.qmd
R2 addressing last minor edit requested by reviewer #1: opening statement on RWE using claims and EHR and linkages; also: upgraded to R 4.3.2

Administrative health insurance claims databases have traditionally been the foundation for pharmacoepidemiologic studies. However, a notable drawback lies in their inability to capture detailed clinical nuances. To overcome this limitation, substantial initiatives are underway, for instance in the FDA Sentinel initiative [@Desai2021], linking claims databases and electronic health records (EHR) to generate real-world evidence (RWE) and complement data from randomized controlled trials (RCTs) [@fdaRWE2018; @Desai2021]. Due to their capture of clinical details, EHR can significantly improve the ability to mitigate imbalances in prognostic factors between treatment groups, especially when linked to administrative claims databases [@asfaw2022]. At the moment, substantial efforts focusing on the linkage of claims databases and EHR are underway, for instance, in the FDA Sentinel Initiative [@Desai2021]. However, prognostic factors coming from EHR are often only partially observed, posing a challenge to the statistical analysis and potentially leading to bias in treatment effect estimates if not handled appropriately [@gorelick2006; @ayilara2019; @groenwold2020].

>>>>

Administrative health insurance claims databases and electronic health records (EHR) are important data modalities to generate real-world evidence (RWE) if they are found fit-for-purpose for the study question at hand. While administrative health insurance claims databases have traditionally been the backbone for the majority of pharmacoepidemiologic studies, a notable drawback lies in their inability to capture important clinical prognostic factors like vital signs and labs. To overcome this limitation, substantial initiatives are underway, for instance in the FDA Sentinel initiative [@Desai2021], linking claims databases and electronic health records (EHR) to generate real-world evidence (RWE) and complement data from randomized controlled trials (RCTs) [@fdaRWE2018; @Desai2021]. Due to their capture of clinical details, EHR can significantly improve the ability to mitigate imbalances in prognostic factors between treatment groups [@asfaw2022]. At the moment, substantial efforts focusing on the linkage of claims databases and EHR are underway, for instance, in the FDA Sentinel Initiative [@Desai2021]. However, prognostic factors coming from EHR are often only partially observed, posing a challenge to the statistical analysis and potentially leading to bias in treatment effect estimates if not handled appropriately [@gorelick2006; @ayilara2019; @groenwold2020].
